---
title: "Lab5_SpicyBois"
author: "SpicyBois"
date: "9/25/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Load tidyverse
library(tidyverse)

# Load and view the data
milk <- read.csv('state_milk_production.csv')
head(milk)
```

#Finding Trends in the Milk Production Data Set


##Distribution of the Yearly Milk Production for 1998 (Trent)
```{r}
milk <- milk %>%
  mutate(milk_million = milk_produced/1000000)
milk1998 <- milk %>%
  filter(year == 1998)

ggplot(data = milk1998, aes(x = milk_million)) +
  geom_histogram(bins = 15) + 
  theme_light()+
  labs(title = "Distribution estimate of milk produced in 1998 by state", x = "Milk Produced (millions of milk pounds)", y = "Count (states)")
```
```{r}
milkcor98 <- milk %>% select(state, year, milk_million) %>% group_by(state) %>%
  summarise(Average = mean(milk_million), Median = median(milk_million)) %>%
  as_tibble()
milkcor98
```

##Yearly Time Series for the Milk Production for Wisconsin (Trent)
```{r}
milkWIS <- milk %>%
  filter(state %in% c('Wisconsin')) %>%
  select(state, year, milk_million)

ggplot(data = milkWIS, aes(x = year, y = milk_million, color = state)) +
  geom_point() + 
  ggtitle('Pounds of milk over time in Wisconsin') +
  geom_smooth(method= "lm", na.rm = TRUE, se = FALSE) +
  xlab('Year') +
  ylab('Milk Produced (Million lb)')
```
**Findings: ** This plot displays some interesting findings, predominantly the decline in milk production from the late 80's to the mid 2000's, followed by a sharp increase of milk production that we currently see. One of Wisconsin's biggest exports is dairy products, so it is slightly puzzling for there to be such a significant dip.

## Team Report: 

**Trent:** I created the plot for the milk production in Wisconsin, as well as the data set for my birth year, 1998. I used the filter function to isolate just the 1998 data, and used a histogram to display the milk produced between all states. For the Wisconsin data, I used the filter function to isolate just Wisconsin's data, and used the geometric point function to graph the findings.
