<<<<<<< Updated upstream
---
title: "Lab7"
author: "SpicyBois"
date: "10/8/2019"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
powerData <- read_delim(file = "household_power_consumption.txt", delim = ";", na = c('', '?'), col_types = cols(
  Date = col_date(format = "%d/%m/%Y"),
  Time = col_time(format = "%H:%M:%S"),
  Global_active_power = col_double(),
  Global_reactive_power = col_double(),
  Voltage = col_double(),
  Global_intensity = col_double(),
  Sub_metering_1 = col_double(),
  Sub_metering_2 = col_double(),
  Sub_metering_3 = col_double()
  ))
powerData <- separate(powerData, Date, into = c("Year", "Month", "Day"), sep = "-", convert = TRUE)
powerData <- separate(powerData, Time, into = c("Hour", "Minute", "Second"), sep = ":", convert = TRUE)
powerData <- powerData %>% mutate(dateTime = make_datetime(Year, Month, Day, Hour, Minute, Second))
```

# Lab 7
**Description of the dataset:** The dataset is details the energy consumtion of a family between 2006-2010. There are 9 varibles including date, time, active power, reactive power, voltage, intensity, active energy for kitchen, active energy for laundry and refrigerator, and active energy for the water-heater and air-conditioner. As for the missing data, we just filtered it out. 

## Why is it that July and Augest the energy graphs look significently different from the other months? (Logan)
**How I came up with this Question:** Loooking at the graph below, which outlines the average Hourly + Minutly consumstion of engery split by month. Each month looks basically looks the same, with the same peeks and drops, execpt for July and Augest. Thus I was curious as to why July and Augest is so drasically diffent energy consumtions from the other months.

```{r}
powerData_logan <- powerData %>%
  filter(!is.na(Sub_metering_1)) %>%
  group_by(Month, Hour, Minute) %>%
  summarise(avg_1 = mean(Sub_metering_1), avg_2 = mean(Sub_metering_2), avg_3 = mean(Sub_metering_3), avg_total = (avg_1 + avg_2 + avg_3))

ggplot(powerData_logan, aes(x = (Hour + (Minute/60)))) + 
  facet_wrap(.~Month) + 
  geom_line(aes(y = avg_1, color = "kitchen")) +
  geom_line(aes(y = avg_2, color = "laundry + refrigerator")) + 
  geom_line(aes(y = avg_3, color = "water-heater + air-conditioner")) + 
  geom_line(aes(y = avg_total, color = "total")) +
  labs(title = "Average Hourly/Minutly energy consumption splip by months and meater", x = "Time of Day", y = "Energy Consummed (watt-hours/Minute)")
```

**Why is this question intersting:** I would think do to the different seasons, each month should show drastically different energy consumtions, with winter and summer showing higher energy consumtions, and fall and spring, having lesser energy consumtions. However, dispite what I thought, there was very minimal change in energy consumtions (as seen in graph bellow, only about ±2 watt-hours). But in July and Augest there is a signifanct change in energy consumtion which is weird thus interesting. 

**Findings:** The graph bellow shows the average energy consumtion for each day of a certain month. It is seen that although there are differeneces between each month that were less aperanct on the graph above, they are slight compared to the giant drop that is seen in July and Augest. It also seen that the drop is mostly due to the meter 1 and 3 which are air-conditional, water heater, and kitchen appliances. Where as meater_2 which landray and refigerator only has slight drop. This concludes that the drop in energy consumtions is due to the lack of use of the air-conditioner, water heater, and kitchen, which most likly means that the family were not home to use these appliances. Since it can be concluded that they were not home plus that fact that it was during summer months, it can be determined that the signifanct change in energy consumtion in July and Augest is most likly do to the family going on vaction.

```{r}
powerData_logan <- powerData %>%
  filter(!is.na(Sub_metering_1)) %>%
  group_by(Month) %>%
  summarise(avg_1 = mean(Sub_metering_1), avg_2 = mean(Sub_metering_2), avg_3 = mean(Sub_metering_3), avg_total = (avg_1 + avg_2 + avg_3))
ggplot(powerData_logan, aes(x = Month)) + 
  geom_line(aes(y = avg_total, color = "total")) + 
  geom_line(aes(y = avg_1, color = "kitchen")) + 
  geom_line(aes(y = avg_2, color = "landray + refrigerator")) + 
  geom_line(aes(y = avg_3, color = "water-heater + air-conditioner")) +
  labs(title = "Average Daily energy consumption grouped by month split by meater", x = "Month", y = "Energy Consummed (watt-hours/day)")
```

**Who Benefits from this information** The energy commpany will most likley benifit from this information, as they can plan ahead, and create less energy for these months.

**Who is harmed from this information** The family is harmed, although they most likly agreed to sharing the energy consumtion energy, they probably did not think that alot of information, such as the times when the family is on vaction, could be found in the data. This means that the family has had the privicy invaded.


##How does energy consumption from kitchen appliances change throughout a day(Tyler)
**How I Came Up With This Question**I wanted to see how energy consumption changed in a day so I chose to focus on one part of the house and I like the kitchen so I chose the Kitchen.
```{r}
dataDay <- powerData %>% dplyr::filter(day == 10, month == 10, year == 2010)
ggplot(dataDay) + geom_smooth(aes(x = hour, y = Sub_metering_1), se = FALSE) + labs(title = "Power Used in the Kitchen in one Day", x = "Hour of Day", y = "Power Used in Kitchen (Watts)")

dinnerEnergy <- dataDay %>% filter(hour <= 21 & hour >= 19) %>% group_by(Date) %>% mutate(total = sum(Sub_metering_1)) %>%  filter(hour ==  19, minute == 0)
totalEnergy <- dataDay %>% group_by(Date) %>% mutate(total = sum(Sub_metering_1)) %>%  filter(hour ==  19, minute == 0)


dinnerEnergy$total/totalEnergy$total * 100
```
**Why is it Interesting** This is interesting cause it shows a change over time for a part of the houses use of energy.

**Findings**This shows that a lot of energy is used mostly durring the dinner hours. Specifically 57% of all of the energy used in the kitchen durring the day is used durring dinner hours(7pm - 9pm). The humps in the graph are when appliances are being used and they tend to occur durring hours of the day when someone would eat food.

**Who Is Benifited and Who is Harmed** An Energy company would benifit as they could see when the energy in the kitchen is being used and can plan to make more money. The Family would be harmed for this as they can see how much energy is used by the kitchen without realizing how much it costs.
