---
title: "Lab 9"
author: "Spicy Bois"
date: "10/17/2019"
output: html_document
---
# How do factors like professor rank, discipline, years since phd, years of service and sex effect the salaries of professors?

### Why is it important/interesting
This question could help find trends and patterns within the salaries of professors to ensure that fair compensation is being provided, especially considering factors such as gender and seniority.

### Who is hiring us?
For the purpose of our the data we found, we are being asked by a college ethics board to analyze the salaries of several different types of professors to ensure equality and fairness in their compensations.

### Description of the Dataset
This Data set has 397 observations (professors) on the following 6 variables: rank, discipline, years since PHD, years of service, gender, and salary, from the accademic year 2008-09.

  * **rank:** Associate Professor (AssocProf), Asstistant Proffessor  (AsstProf), Proffessor (Prof)

  * **discipline:** a factor with levels A (“theoretical” departments) or B (“applied” departments).

  * **yrs.since.phd:** The years since professor got PhD.

  * **yrs.service:** The years years the professor has worked/service.

  * **sex:** Female or Male

  * **salary:** nine-month salary, in dollars.

Data Source: https://vincentarelbundock.github.io/Rdatasets/datasets.html

Codebook Link: https://vincentarelbundock.github.io/Rdatasets/doc/carData/Salaries.html

## Individual Questions

### How does the rank of a professor impact their salary? (Jack)
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

##How do salaries of professors vary based on their rank? (Jack)
```{r}

dataBase <- read_csv("Salaries.csv")

Salaries <- dataBase %>% select(rank, salary)

ggplot(data = Salaries, mapping = aes(x = rank, y = salary)) + geom_boxplot() + theme_bw()

```
**Findings** : As expected, as the rank of the professor increases, so does the salary. The IQR of each rank is increases as the rank increases. However, professors have the largest range of values, some being even below both rankings. There are also outliers for the rank of professor on the top end of the range. Although, for the most part, as rank increases so does salary, there are values that contradict this for all ranks.

<<<<<<< HEAD
### How does the seniority of a professor impact their salary? (Trent)
```{r}
library(tidyverse)

data <- read_csv("Salaries.csv")
seniority <- data %>% select(yrs.service, salary)

ggplot(data) + geom_point(aes(x = yrs.service, y = salary)) + geom_smooth(aes(x = yrs.service, y = salary, color = 'Trend line')) + labs(title = "Distribution of Salaries by Years of Service", y = "Annual Salary (US Dollars)", x = "Years of Service (years)") + facet_grid(rank~ .) + theme_bw()
```

### Which discipline makes the most money? (Logan)

### How does the gender of a professor impact their salary? (Joe)

### How does the amount of time a teacher has had their Phd effect their Slary? (Tyler)

## Team Report
**Trent:**

**Logan:**

**Tyler**

**Jack**For my individual graph, I did how salaries of professors vary based on their rank. The data was already tidy so I just selected the columns I needed using the select function. For my graph, I used a box plot because it shows the salaries compared to the ranking the best. I kept it simple and just used the boxplot function from ggplot as well as a theme to change the graph background.

**Joe**
=======
