---
title: "Lab 9"
author: "Spicy Bois"
date: "10/17/2019"
output: html_document
---
# How do factors like professor rank, discipline, years since phd, years of service and sex effect the salaries of professors?

### Why is it important/interesting
This question could help find trends and patterns within the salaries of professors to ensure that fair compensation is being provided, especially considering factors such as gender and seniority.

### Who is hiring us?
For the purpose of our the data we found, we are being asked by a college ethics board to analyze the salaries of several different types of professors to ensure equality and fairness in their compensations.

### Description of the Dataset
This Data set has 397 observations (professors) on the following 6 variables: rank, discipline, years since PHD, years of service, gender, and salary, from the accademic year 2008-09.

  * **rank:** Associate Professor (AssocProf), Asstistant Proffessor  (AsstProf), Proffessor (Prof)

  * **discipline:** a factor with levels A (“theoretical” departments) or B (“applied” departments).

  * **yrs.since.phd:** The years since professor got PhD.

  * **yrs.service:** The years years the professor has worked/service.

  * **sex:** Female or Male

  * **salary:** nine-month salary, in dollars.

Data Source: https://vincentarelbundock.github.io/Rdatasets/datasets.html

Codebook Link: https://vincentarelbundock.github.io/Rdatasets/doc/carData/Salaries.html

## Individual Questions

### How does the rank of a professor impact their salary? (Jack)
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

##jack(edit)
dataBase <- read_csv("Salaries.csv")


Salaries <- dataBase %>% select(rank, salary)

ggplot(data = Salaries, mapping = aes(x = rank, y = salary)) + geom_boxplot()


```

### How does the seniority of a professor impact their salary? (Trent)
```{r}
library(tidyverse)

data <- read_csv("Salaries.csv")
seniority <- data %>% select(yrs.service, salary)

ggplot(data) + geom_point(aes(x = yrs.service, y = salary)) + geom_smooth(aes(x = yrs.service, y = salary, color = 'Trend line')) + labs(title = "Distribution of Salaries by Years of Service", y = "Annual Salary (US Dollars)", x = "Years of Service (years)") + facet_grid(rank~ .) + theme_bw()
```

### Which discipline makes the most money? (Logan)

### How does the gender of a professor impact their salary? (Joe)

### How does the amount of time a teacher has had their Phd effect their Slary? (Tyler)

## Team Report
**Trent:**

**Logan:**

**Tyler**

**Jack**

**Joe**
