---
title: "Lab3"
author: "SpicyBois: Jack, Joe, Logan, Trent, Tyler"
date: "9/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab 3
```{r}
library(tidyverse)
source("http://www.openintro.org/stat/data/cdc.R")
cdc <- as_tibble(cdc)
```

## TEAM QUESTION: Which attributes effect how healthy someone is/feels?
### Why is the Question important/interesting?
Finding the different attributes that effect how someone views their health is important for varies reasons. One of which, is to help the general population be/feel healthyer. If we know that most of the people who have smoke over 100 cigerates typically have a poor health, then we can conclude that smoking causes health to decrease. Also we can find certian attributes that do not effect the general health of someone. 

### Does smoking effect general health? (TEAM)
```{r}
smoke_data <- cdc %>%
  group_by(genhlth) %>%
  summarize(count = mean(smoke100))

ggplot(smoke_data, aes(x = genhlth, y = count, fill = genhlth)) + #X axis -> General Health, Y axis -> Mean Age
  geom_col() + #Display a bar Graph

  #Edit visuals
  theme_dark() + 
  
  #Add Text
  labs(title = "Smoking vs. General Health.", x = "General Health", y  = "Average smokers (0 - No, 1 - Yes)") + 
  scale_fill_discrete(name = "Health", labels = c("Excellent", "Very Good", "Good", "Fair", "Poor")) # Edit Legend
```

### Does whether or not people exercised effect how they describe their health? (Jack)
```{r}
library(tidyverse)
source("http://www.openintro.org/stat/data/cdc.R")
cdc <- as_tibble(cdc)

ggplot(data = cdc, (mapping =  aes(x = exerany))) +
  stat_count()+
  facet_grid(.~genhlth)+
  labs(title = "Whether People Excercised in the Past Month, and How They Describe Their Health")
  
```
**Findings:** The plot shows that although many people worked out, there were still people who worked out and stated that their health was not very good. On the other hand, there were people who did not work out that stated their health was "excellent." All in all, there were people who worked out and didn't work out, but it seems their view on their own health seemed to vary despite working out or not.


### Does desired weight effect general health? (Tyler)
```{r}
library(tidyverse)
source("http://www.openintro.org/stat/data/cdc.R")
cdc <- as_tibble(cdc)

ggplot(data = cdc, (mapping = aes(x = genhlth, y = wtdesire, color = gender,))) +
  geom_point() +
  labs(title = "General Health  VS  Desired Weight") + 
  theme_dark() + 
  facet_grid(gender~.)
```

**Findings:** This plot shows that There isn't a difference in someones general health and their desired weight. 

## Does having health insurance effect General Health? (Trent)
```{r}
library(tidyverse)
source("http://www.openintro.org/stat/data/cdc.R")
cdc <- as_tibble(cdc)

ggplot(cdc, aes(genhlth, ..count..)) + geom_bar(aes(color = hlthplan), position = "dodge")+
  labs(title = "Amount of People who have a Healthplan, and How they Descibe their Health") + 
  facet_grid(gender~.)
```

**Findings:** This plot shows how people with good - excellent health tend to have a healthplan care, where as people who say they have fair-good health tend to not have a plan. It it interesting to note that people who descibe their health as "excellent" have less healthcare plans than people who say they have "very good" health. This may be due to some people have such excellent health that they do not feel they need a healthplan. No difference between genders.


### Does gender effect general health? (Logan)
```{r}
ggplot(cdc, aes(x = genhlth, y = gender)) + 
  stat_bin2d() + 
  
  #Edit visuals
  theme_dark() + 
  
  #Add Labels
  labs(title = "Gender vs. General Health.", x = "General Health", y = "Gender", caption = "m - Male and f - Female") + 
  scale_fill_continuous(name = "Number of People", limits=c(0, 4000), breaks=seq(0,4000,by=1000)) # Edit Legend
```

**Findings:** The graph shows that the total count of males and females in each catagory of General Health is not significant different. This demonstates that gender does not play a major rule in effecting health. 


### What connection is there between General Health and the weight of a person proportional to their height? (Joe)
```{r}
library(tidyverse)
cdc <- as_tibble(cdc)

ggplot(data=cdc, (mapping = aes(x= height, y= weight, color = genhlth)))+ 
  geom_smooth(se = FALSE) +
  labs(title = "Weight vs Height Divided into Catagories Based on Self-Reported Health") + 
  labs(caption = "The uncolored line is a general representation of a healthy correlation between height and weight.") +
  theme_light() +
  facet_wrap(genhlth~.) +
  geom_abline(intercept= -60, slope= 3.33)
```

**Findings:** As people rate themselves lower on the general health scale, the percentage of the population that is considered overweight increases significantly.

### Conclustion
As a team, we chose to graph the relationship between people's general health and wheter or not they smoke, as smoking is what we considered to be the unhealthiest factor from the dataset. Unsuprisingly, people who claim to have "poor" health is also the group with the most smokers. The other most significant factor that effected people's overall sense of well-being was their weight. The percentage of the population that is considered overweight tended to indicate fair-poor health quality. We also discovered that some factors like gender have no effect on how someone feels about their health. Overall, people who feel better about their health tend to be not-overweight, non-smokers, and have some sort of healthcare plan. So our reccomdation to the public would be to eat health, exercise, and reduce/stop smoking in order to feel as healthy as possible.

### Team Contributions
We each worked on our individual graphs, Logan made the TEAM chart, and Trent worked on the conclusion