---
title: "Lab3"
author: "SpicyBois: Jack, Joe, Logan, Trent, Tyler"
date: "9/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab 3
```{r}
library(tidyverse)
source("http://www.openintro.org/stat/data/cdc.R")
cdc <- as_tibble(cdc)
```

## Which attributes effect how healthy someone is/feels?
### Why is the Question important/interesting?
Finding the different attributes that effect how someone views their health is important for varies reasons. One of which, is to help the general population be/feel healthyer. If we know that most of the people who have smoke over 100 cigerates typically have a poor health, then we can conclude that smoking causes health to decrease. Also we can find certian attributes that do not effect the general health of someone. 

### Does smoking effect general health? (TEAM)
```{r}
smoke_data <- cdc %>%
  group_by(genhlth) %>%
  summarize(count = mean(smoke100))

ggplot(smoke_data, aes(x = genhlth, y = count, fill = genhlth)) + #X axis -> General Health, Y axis -> Mean Age
  geom_col() + #Display a bar Graph

  #Edit visuals
  theme_dark() + 
  
  #Add Text
  labs(title = "Smoking vs. General Health.", x = "General Health", y  = "Average smokers (0 - No, 1 - Yes)") + 
  scale_fill_discrete(name = "Health", labels = c("Excellent", "Very Good", "Good", "Fair", "Poor")) # Edit Legend
```

**Findings:** 


### Does desired weight effect general health? (Tyler)
```{r}
library(tidyverse)
source("http://www.openintro.org/stat/data/cdc.R")
cdc <- as_tibble(cdc)

ggplot(data = cdc, (mapping = aes(x = genhlth, y = wtdesire, color = gender,))) +
  geom_point() +
  labs(title = "General Health  VS  Desired Weight") + 
  theme_dark() + 
  facet_grid(gender~.)
```

**Findings:** This plot shows that There isn't a difference in someones general health and their desired weight. 

## Does having health insurance effect General Health? (Trent)
```{r}
library(tidyverse)
source("http://www.openintro.org/stat/data/cdc.R")
cdc <- as_tibble(cdc)

ggplot(cdc, aes(genhlth, ..count..)) + geom_bar(aes(color = hlthplan), position = "dodge")+
  labs(title = "Amount of People who have a Healthplan, and How they Descibe their Health") + 
  facet_grid(gender~.)
```

**Findings:** This plot shows how people with good - excellent health tend to have a healthplan care, where as people who say they have fair-good health tend to not have a plan. It it interesting to note that people who descibe their health as "excellent" have less healthcare plans than people who say they have "very good" health. This may be due to some people have such excellent health that they do not feel they need a healthplan. No difference between genders.


### Does gender effect general health? (Logan)
```{r}
ggplot(cdc, aes(x = genhlth, y = gender)) + 
  stat_bin2d() + 
  
  #Edit visuals
  theme_dark() + 
  
  #Add Labels
  labs(title = "Gender vs. General Health.", x = "General Health", y = "Gender", caption = "m - Male and f - Female") + 
  scale_fill_continuous(name = "Number of People", limits=c(0, 4000), breaks=seq(0,4000,by=1000)) # Edit Legend
```

**Findings:** The graph shows that the total count of males and females in each catagory of General Health is not significant different. This demonstates that gender does not play a major rule in effecting health. 

### Conclustion:

### Team Report