---
title: "Lab 15 - SpicyBois"
author: "Trent Collins"
date: "12/13/2019"
output: html_document
---
# Lab 15 - It's Showtime!

## Team Question - How do different factors affect the overall quality and financial success of a film?

**Domain Expert:** Movie studios and executives who have to invest big money into the films they produce

**Data:** We used data from Data World
link: https://data.world/crowdflower/blockbuster-database

**Why is it important/interesting?** There are several different factors presented in our data set: date of release, run time, studio, genre, etc. All of these factors have a huge impact to how a movie may be perceived, and how financially successful it will be. Determining which of these factors do and do not have an impact on the financial and critical success of a movie would be very enlightening to people who work in the film industry to know what factors can lead to a hit.

**Findings:** Overall, the most ideal movie based on our findings for each categorical variable would be one that is over 2 hours long, released by a large studio such as Warner Bros/Paramount/Fox/Disney/Universal, have a summer release (May, June , July), and would be in the history genre. 

**Recommendations:** While there are certain factors that seem to play better with critics and audiences, and make movies more money, the most important thing a filmmaker can focus on is making the best possible version of their vision. So while this data can help executives and produces decide on whether they should invest in a film, it should not affect how films are being made or impact the vision of a filmmaker. 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(purrr)
library(modelr)

movies <- read_csv("movies.csv") %>% select(-c("poster_url", "genres")) %>% separate(release_date, sep="-", into=c("release_day", "release_month", "release_year")) %>% mutate(release_day = as.numeric(release_day), release_year = as.numeric(release_year))
```

## Individual Sections

### Trent Collins - How does the length of a film effect critic scores, audience scores, and box-office gross?

**Why is it important/interesting?** The lenghth of a film is something that a lot of people pay attention to, and can effect people's perception of the film, ie being too long or too short. Knowing how the runtime can effect audience/critic scores would be helpful to film studios and directors so that they know how to pace and schedule their films.

```{r}
model <- lm(length ~ audience_freshness, data = movies)

grid <- movies %>%
  data_grid(audience_freshness)

grid <- grid %>% 
  add_predictions(model)

ggplot(data = movies, mapping = aes(x = audience_freshness,  y = length)) + 
  geom_point(mapping = aes(color = length), na.rm = TRUE) +
  geom_line(aes(y = pred), data = grid, color = "red", size = 1) +
  labs(title = "How the Runtime of a Film effects the Audience Score", x = "Audience Score (1-100)", y = "Runtime (minutes)")
```

```{r}
model <- lm(length ~ rt_freshness, data = movies)

grid <- movies %>%
  data_grid(rt_freshness)

grid <- grid %>% 
  add_predictions(model)

ggplot(data = movies, mapping = aes(x = rt_freshness,  y = length)) + 
  geom_point(mapping = aes(color = length), na.rm = TRUE) +
  geom_line(aes(y = pred), data = grid, color = "red", size = 1) +
  labs(title = "How the Runtime of a Film effects the Critics Score", x = "Critics Score (1-100)", y = "Runtime (minutes)")
```

```{r}
model <- lm(length ~ adjusted, data = movies)

grid <- movies %>%
  data_grid(adjusted)

grid <- grid %>% 
  add_predictions(model)

ggplot(data = movies, mapping = aes(x = adjusted,  y = length)) + 
  geom_point(mapping = aes(color = length), na.rm = TRUE) +
  scale_x_continuous(breaks=seq(0,3000000000,300000000)) +
  geom_line(aes(y = pred), data = grid, color = "red", size = 1) +
  geom_hline(yintercept = 119.25, color = "black") +
  labs(title = "How the Runtime of a Film effects the Adjusted Gross of the Film's Boxoffice", x = "Adjusted Gross (American Dollars)", y = "Runtime (minutes)", caption = "Black Line represents a runtime of roughy 119 minutes (average length of a film from 1975-2015)")
```

**New tools used:** I used the lm, grid, and prediction tools to graph predictions and find trends from my data

**Findings:** The mean movie length is just over 119 minutes. In terms of audience scores and critic scores, the rating of a movie is slightly higher if it is longer, however the length does not have a significant impact on how they are rated. Adjusted gross however is affected by the runtime of a film, as show in the linear trend of the data set, longer movies tend to make more money. This could be due to the big summer blockbusters, such as Marvel films and Star Wars films making most of the money, as well as tending to run over two hours long.

**How does this relate to the overall question?** Clearly, while in most cases the quality of a film is not significantly affected by runtime, movies tend to do better financially when they are longer (over two hours).

**Ethical Implications:** While I personally cannot recommend that studios make their movies longer based on the results of the data because I think the length of most movies are too long as it is now-a-days, there is a clear trend from the data that longer movies just do better. Knowing that, studios should not press to make movies longer, and instead just let the directors/screen writers tell their stories in the best way they can, and let the runtimes just happen naturally, because if studios enforce longer runtimes, that can lead to movie-makers just adding irrelevant garbage to pad the run time, and no one wants that.


### Logan Richards - How does the size of the Studio effect the quality and success of a movie? 

**Why is it important/interesting?** It is interesting, to understand whether the bigger studios make better movies, or do they just creating a bunch of movies hopping 1-2 are great.

```{r}
rename_studio <- function(names) {
  returnVec <- vector("character", length = length(names))
  
  for(i in 1:length(names)) {
    studio_name <- names[i]
    
		if(studio_name == "20thCentury-Fox" | studio_name == "20thCentury-FoxFilmCorporation" | studio_name == "20thCenturyFox" | studio_name == "20thCenturyFoxFilmCorporation" | studio_name == "Fox" | studio_name == "FoxSearchlightPictures") {
			returnVec[i] = "20thCenturyFox"
		} else if(studio_name == "Carolco" | studio_name == "CarolcoPictures") {
			returnVec[i] = "Carolco"
		} else if(studio_name == "Columbia" | studio_name == "ColumbiaPictures") {
			returnVec[i] = "Columbia"
		} else if(studio_name == "DreamWorks" | studio_name == "DreamWorksPictures") {
			returnVec[i] = "DreamWorks"
		} else if(studio_name == "Imagine" | studio_name == "ImagineEntertainment") {
			returnVec[i] = "Imagine"
		} else if(studio_name == "Lionsgate" | studio_name == "LionsgateFilms") {
			returnVec[i] = "Lionsgate"
		} else if(studio_name == "Marvel" | studio_name == "MarvelStudios") {
			returnVec[i] = "Marvel"
		} else if(studio_name == "Metro-Goldwyn-Mayer" | studio_name == "MGM") {
			returnVec[i] = "MGM"
		} else if(studio_name == "NewLine" | studio_name == "NewLineCinema") {
			returnVec[i] = "NewLine"
		} else if(studio_name == "Paramount" | studio_name == "ParamountPictures") {
			returnVec[i] = "Paramount"
		} else if(studio_name == "Summit" | studio_name == "SummitEntertainment") {
			returnVec[i] = "Summit"
		} else if(studio_name == "Touchstone" | studio_name == "TouchstonePictures") {
			returnVec[i] = "Touchstone"
		} else if(studio_name == "UnitedArtists" | studio_name == "UnitedArtistsPictures" | studio_name == "UnitedFilmDistributionCompany") {
			returnVec[i] = "UnitedArtists"
		} else if(studio_name == "Universal" | studio_name == "UniversalPictures" | studio_name == "UniversalStudios") {
			returnVec[i] = "Universal"
		} else if(studio_name == "Disney" | studio_name == "WaltDisneyPictures" | studio_name == "WaltDisneyProductions") {
			returnVec[i] = "Disney"
		} else if(studio_name == "WarnerBros" | studio_name == "WarnerBros." | studio_name == "WarnerBros.Pictures") {
			returnVec[i] = "WarnerBros"
		} else {
			returnVec[i] = studio_name
		}
  }
  
  return(returnVec)
}
```

```{r, message=FALSE, warning=FALSE}
movies_studio <- movies %>% 
  separate(studio, sep = "/", into = c("main_studio", "sub_studio_1", "sub_studio_2")) %>% 
  gather(main_studio, sub_studio_1, sub_studio_2, key = "studio_rank", value = "studio") %>%
  filter(!is.na(studio)) %>%
  mutate(studio = rename_studio(gsub(" ", "", studio))) %>%
  arrange(title)
```

```{r}
movies_group <- movies_studio %>% 
  group_by(studio) %>% 
  summarise(sum_gross = sum(adjusted), sum_rating = sum(imdb_rating), count = n(), avg_per_movie = (sum_gross/count), avg_rating_per_movie = (sum_rating/count)) %>%
  arrange(studio)

ggplot(movies_group, aes(x = reorder(studio, -avg_per_movie))) + 
  geom_col(aes(y = avg_per_movie)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  labs(title = "Average Gross per movie per Studio", x = "Studio", y = "Average Gross per Movie")

ggplot(movies_group, aes(x = reorder(studio, -avg_rating_per_movie))) + 
  geom_col(aes(y = avg_rating_per_movie)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  labs(title = "Average IMDB Rating per movie per Studio", x = "Studio", y = "Average IMDB Rating per Movie")

ggplot(movies_group, aes(x = reorder(studio, -count))) + 
  geom_col(aes(y = count)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  labs(title = "Number of Movies made per Studio", x = "Studio", y = "Number of Movies")

ggplot(movies_group, aes(x = reorder(studio, -sum_gross))) + 
  geom_col(aes(y = sum_gross)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  labs(title = "Gross Sum per Studio", x = "Studio", y = "Gross Sum")
```

**New tools used:** I used theme to rotate the x axis to better view the studio names. I created a function in order to better tidy the studio names, because that was a mess. 

**Findings:** To determine the quality I looked at both gross and IMDB ratings. To determine the scale of the studio I looked at the number of movies created. 
I determined that Mid size studios seem to create better quality movies. If you look at the largest studio, WarnerBros, it is the 21st studio of average gross and the 23rd in ratings. However, the studio that has the best gross, Marvel, is only 14th largest studio. Whereas Lightstorm has the average best scores yet it is near the bottom of the studio sizes. Looking over the 2nd-10th place one’s similar trends continue, the studios that are higher quality, tend to be mid-size studios.

**How does this relate to the overall question?** The Studio does impact the quality of the movie.

**Ethical Implications:** It is seen that large size studios tend to be in the middle of the ranks for both gross and ratings. Yet they are still at the top for the total gross made. This means that despite creating lesser quality movies, they can still make huge amounts of money. The ethical implications of my findings are that large size studios might care more about quickly making movies more than creating a quality movie.

## Tyler Bischoff - Hows does the main genre of a movie effect its rating and gross amount of money made?

**Why is this interesting?** This is an important and interesting question because it can show how movies do in the box office based on the genre. This can show how mush a movie could be projected to make depending on their genre

```{r}

ggplot(data = movies) +
  geom_bar(stat = "identity", mapping = aes(x = genre_1, y = adjusted)) +
  theme_bw()+
  theme(axis.text.x = element_text(angle = -60, hjust = 1)) +
  labs(title = "How Gross Worldwide Money Made is effected by Genre", x = "Genre", y = "Combinded Gross Money Made")

ggplot(data = movies) +
  geom_bar(stat = "identity", mapping = aes(x = genre_1, y = rt_freshness)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = -60, hjust = 1)) +
  labs(title = "How ratings are effected by genre", x = "Genre", y = "Combined Rotten Tomatoes Freshness Score")


```

**New tools used:** For this I used GGPLOT2, importing data, and tidying data.

**Findings:** I have found that history movies have made the most money overall and have a high total freshness score. However, war movies have the highest combines rotten tomatoes score. This adds to the overall team question as it can show how certain genres of movies can make more money and do betetr in the box office. For this I used GGPLOT2, importing data, and tidying data. 

**Ethical Implications:** This can help movie producers and studios figure out what genre they can make a movie to make a lot of revenue


### Joe Lawton - Does the quality or target audience of a movie effect its box office returns?

**Why is it important/interesting?** My question is interesting because it shows that investing resources and passion into making a good movie usually rewards everyone at the box office. The audience is happy, the critics are happy, and the production company is hauling in the cash. Making the best movie possible maximizes social surplus.

```{r}
blockbuster <- movies %>% select('title', 'adjusted', 'audience_freshness', 'rt_freshness') %>% filter(audience_freshness > 90) %>% filter(rt_freshness < 90) %>% arrange(desc(adjusted)) %>% slice(1:5) %>% mutate(stringr::str_trunc(title, 20))
colnames(blockbuster)[5] <- "newname"
bavg <- blockbuster %>% select('adjusted') %>% map(mean)
bavg

oscarbait <- movies %>% select('title', 'adjusted', 'rt_freshness', 'audience_freshness') %>% filter(rt_freshness > 90) %>% filter(audience_freshness < 80) %>% arrange(desc(adjusted)) %>% slice(1:5) %>% mutate(stringr::str_trunc(title, 20))
colnames(oscarbait)[5] <- "newname"
oavg <- oscarbait %>% select('adjusted') %>% map(mean)
oavg

excellence <- movies %>% select('title', 'adjusted', 'audience_freshness', 'rt_freshness') %>% filter(audience_freshness > 90) %>% filter(rt_freshness > 90) %>% arrange(desc(adjusted)) %>% slice(1:6) %>% mutate(stringr::str_trunc(title, 20))
colnames(excellence)[5] <- "newname"
eavg <- excellence %>% select('adjusted') %>% map(mean)
eavg

trash <- movies %>% select('title', 'adjusted', 'audience_freshness', 'rt_freshness') %>% filter(audience_freshness < 50) %>% filter(rt_freshness < 50) %>% arrange(desc(adjusted)) %>% slice(1:5) %>% mutate(stringr::str_trunc(title, 20))
tavg <- trash %>% select('adjusted') %>% map(mean)
tavg

ggplot()+
  geom_col(data = blockbuster, mapping = aes(x = newname, y = adjusted))+
  labs(title = "Top 5 Grossing Crowdpleasers", x = "Movie Title", y = "Box Office Gross")+
  geom_hline(yintercept = 867308679)
ggplot()+
  geom_col(data = oscarbait, mapping = aes(x = newname, y = adjusted))+
  labs(title = "Top 5 Grossing Critic Pleasers", x = "Movie Title", y = "Box Office Gross")+
  geom_hline(yintercept = 764569664)
ggplot()+
  geom_col(data = excellence, mapping = aes(x = newname, y = adjusted))+
  labs(title = "Top 5 Grossing Movies Loved by Audience and Critics", x = "Movie Title", y = "Box Office Gross")+
  geom_hline(yintercept = 1188399713)
ggplot()+
  geom_col(data = trash, mapping = aes(x = title, y = adjusted))+
  labs(title = "Top 5 Grossing Disliked Movies", x = "Movie Title", y = "Box Office Gross")+
  geom_hline(yintercept = 622350515)
```

**New tools used:** I used the mapping tool from purrr to grab the average of a whole collumn of values from each tibble. 

**Findings:** Movies that are well liked by audience and critics to far better than any other category of movie. Movies that only critics like seem to do the worst. This lets production companies know that it is very much so worth it to create the best movie possible. If this were obeyed by every production studio in Hollywood, more high quality movies would be made instead of movies that are attempting to grab a quick profit.

**Ethical Implications:** I believe that this shows that movie studios in Hollywood are not only unethical with their systematic destruction of established titles, but they're not even being efficient with there money making. They're just being lazy


### Jack Braglia - How does the month of the year affect how a movie is rated and the gross amount of money they make?

**Why is it important/interesting?** This question is important because it shows what months will yield the best ratings and the most money. It is interesting because we are able to visualize the release months and how movies fair with it.

```{r}
ggplot(data = movies) +
  geom_bar(stat = "identity", mapping = aes(x = release_month, y = adjusted, color = release_month), fill="black") +
  theme_classic()+
  theme(axis.text.x = element_text(angle = -60, hjust = 1)) +
  labs(title = "How Gross Worldwide Money Made is effected by Release Month", x = "Release Month", y = "Combinded Gross Money Made")+
  coord_flip()+
  scale_x_discrete(limits = month.abb)


ggplot(data = movies) +
  geom_bar(stat = "identity", mapping = aes(x = release_month, y = imdb_rating, color = release_month), fill="black") +
  theme_classic()+
  theme(axis.text.x = element_text(angle = -60, hjust = 1)) +
  labs(title = "How ratings are effected by Release Month", x = "Release Month", y = "Combined IMDb Ratings")+
  coord_flip()+
  scale_x_discrete(limits = month.abb)

```

**New tools used:** I utilized ggplot in order to plot my findings. I used coord flip to make a horizontal bar graph. I also have colored lines in the bar graphs in order to show how each movie contributed to the ratings/money made. In order to order the months I utilized scale_x_discrete adding a limit to months in order for them to be in order.

**Findings:** As far as gross money is concerned, the summer months certainly made the most with May being the highest grossing month. You can see with the lines in the bar graph that there seems to be more quality movies (according to the gross money made) in May compared to June, where it seems as if there are a lot of movies in quantity, but a good amount of those movies did not make money. For ratings, a similar trend was found in having summer months being the leader for most rated movies. June is the highest. Another trend for both money made and ratings is that November and December have increased amounts of money made and ratings had. This is likely due to the holiday season and there being more free time.

**How does this relate to the overall question?** This relates to the overall question because it is clear that release month has an impact on the overall success of a movie, whether that be ratings or gross money made. 

**Ethical Implications:** Ethical implications for my findings could be that although a great movie could be created in a month like January, the amount of free time a person has is likely less than that of a holiday season, thus making less money and having less overall ratings.


## Lab 2 Reflection

**Team Relections:** We as a team believe that we all learned R very efficiently, we all had contributions to porjects and insight that helped everyone so that we could succeed in utilizing R/RStudio for data science, adding a great tool to our belts. If there's one thing we could have done better this semester it would be communication.

**Trent Collins:** I feel like my personal goal was met fairly successfully. I learned a lot this semseter about data and using R, which was my main goal. I still need practice on subjects such as Models, but overall I learned lots of useful skills.

**Tyler Bischoff:** I feel like we did good on our goal. We started off not really working together but b lab 9 we started to get together and work on the lab as a team. I feel like this is one this we should keep doing. One thing that we did need to work on was organization through gitKraken just to avoid merge conflicts. I feel like my goals have changed as 6 months after graduation I would like to be going into graduate school. My 5-year goal has stayed the same

**Jack Braglia:** My goals have yet to change, I am now more equipped in the data science field and I am excited for what is to come! I learned how to work with a group effectively, as well as improved communicationn techniques. I accomplished learning R in an effective way, in fact, I've started a model for checking lines on sports as I would one day hope to become a professional sports gambler. If I could give myself advice at the beginning of the semester it would be to keep doing problems to solidify my efficiency in R, to stop missing class, and to start getting stuff done/studying earlier.

**Joe Latwon:** My six-month goal is the same still. I just want to study something that is interesting to me and applicable in a workspace without spending more than six years here. My five year goal has changed. I think that it would be perfectly fine for me to be working an office job, as long as the office is a supportive and comfortable place to work. I learned a lot about coding this semester and even more about working in groups. I realized how much easier work goes when a team works as one. I wish that we had started working as a team earlier on into the year. I would also tell myself to stop staying up so late and to keep asking questions when I needed help.

**Logan Richards:** My goals have stayed the same as stated in Lab 2. I learned more about statitis and the impact it can have on almost everything. If I could give myself advice at the beginning of the semester I would team myself to start trying to communicate with my parterners more often about the lab, stop the tendency to procrastinate on the labs, and continues the overall work effort I did on the labs. 


## Team Reflection

**Trent Collins:** I worked on the overall Lab question section, as well as my individual section as well as overall formatting of the lab

**Jack Braglia:** I worked on my individual subsection and helped with the Lab 2 Reflection

**Joe Latwon:** I worked on my individual subsection and helped with the team question

**Tyler Bischoff:** I helped to find and tidy the data set as well as worked on my individual subsection 

**Logan Richards:** I worked on my individual subsection



