---
title: "Lab 8"
author: "Spicy Bois"
date: "10/17/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
```


# Who is hiring us?
**Domain Expert:** For the purpose of our the data we found, we are being asked by a college ethics board to analyze the salaries of several different types of professors to ensure equality and fairness in their compensations. 

# Overall Question (TEAM)

**Question:** How do factors like professor rank, discipline, years since phd, years of service and sex effect the salaries of professors. 

**Why is it important/interesting:** This question could help find trends and patterens within the salaries of professors to ensure that fair compesnation is being provided, especially considering factors such as gender and senority. 

# Description of the Dataset
This Data set has 397 observations(professors) on the following 6 variables: rank, discipline, years since PHD, years of service, gender, and salary.

# Individual Questions
## Question ()
**Why is it important/interesting:**

# Visualizations
## Challa (Joe)
```{r}
library(tidyverse)
library("readxl")
challa <- read_excel("eastafricalakedata.xlsx", sheet = "Challa", col_names = FALSE, col_types = NULL, na = "", skip = 3)

challa <- rename(challa, `Date` = `...1`, `Varve Number`= `...2`, `VarveThickness` = `...3`, `LightThickness` = `...4`, `DarkThickness`= `...5`)

ggplot(data = challa, mapping = aes(x= Date))+
 geom_smooth(aes(y= LightThickness, color = "LightThickness"))+
  geom_smooth(aes(y=DarkThickness, color = "DarkThickness")) +
  geom_smooth(aes(y= VarveThickness, color="VarveThickness"))+
  labs(title = "Measurement of sediment deposit on the bed of Lake Challa", caption= "Varve is the cumulative value of DarkThickness and LightThickness.", fill= "Layers")+
  xlab("Date (AD)")+
  ylab("Thickness of layer in milimeters")
```

## Tanganyika (Trent)
```{r}
VOI_Rainfall <- read_excel("East Africa lake data.xlsx", sheet = "Tanganyika")
Tanganyika <- select()
```

## Malawi (Tyler)
```{r}

```

## CRU and MAM Rainfall (Jack)
```{r}

```


## VOI Rainfall (Logan)
```{r, message=FALSE}
VOI_Rainfall <- read_excel("East Africa lake data.xlsx", sheet = "VOI rainfall station data")
VOI_Rainfall_tidy <- VOI_Rainfall %>% select(-c(`...14`, `...15`)) %>% 
  gather("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec", key = "Month", value = "Rainfall") %>%
  mutate(Month = if_else(Month == "Jan", 1, if_else(Month == "Feb", 2, if_else(Month == "Mar", 3, if_else(Month == "Apr", 4, if_else(Month == "May", 5, if_else(Month == "Jun", 6, if_else(Month == "Jul", 7, if_else(Month == "Aug", 8, if_else(Month == "Sep", 9, if_else(Month == "Oct", 10, if_else(Month == "Nov", 11, if_else(Month == "Dec", 12, 0))))))))))))) %>% 
  arrange(Year) %>%
  filter(!is.na(Rainfall))

labels <- c(Female = "Women", Male = "Men")

ggplot(VOI_Rainfall_tidy, aes(x = Year, y = Rainfall)) + 
  facet_wrap(.~Month, labeller=labeller(Month =  c(`1` = "January", `2` = "February", `3` = "March", `4` = "April", `5` = "May", `6` = "June", `7` = "July", `8` = "August", `9` = "September", `10` = "October", `11` = "November", `12` = "December"))) + 
  geom_line() + 
  labs(title = "VOI Rainfall for every month vs. year", y = "Rainfall (cm)", x = "Year")
```




