---
title: "Lab 8"
author: "Spicy Bois"
date: "10/17/2019"
output: html_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
```


# Who is hiring us?
**Domain Expert:** For the purpose of our the data we found, we are being asked by a college ethics board to analyze the salaries of several different types of professors to ensure equality and fairness in their compensations. 

# How do factors like professor rank, discipline, years since phd, years of service and sex effect the salaries of professors. (TEAM)
**Why is it important/interesting:** This question could help find trends and patterens within the salaries of professors to ensure that fair compesnation is being provided, especially considering factors such as gender and senority. 

# Description of the Dataset
This Data set has 397 observations (professors) on the following 6 variables: rank, discipline, years since PHD, years of service, gender, and salary, from the accademic year 2008-09.
  * **rank:** Associate Professor (AssocProf), Asstistant Proffessor  (AsstProf), Proffessor (Prof)
  * **discipline:** a factor with levels A (“theoretical” departments) or B (“applied” departments).
  * **yrs.since.phd:** The years since professor got PhD.
  * **yrs.service:** The years years the professor has worked/service.
  * **sex:** Female or Male
  * **salary:** nine-month salary, in dollars.

# Individual Questions
## Which discipline makes the most money? (Logan)
**Why is it important/interesting:** This question is interesting becuase it shows which deperment is valued the most in a typical campus. Which shows things like wether we value Arts more or Science more.

# Visualizations
## Challa (Joe)
```{r, warning=FALSE, message=FALSE}
challa <- read_excel("East Africa lake data.xlsx", sheet = "Challa", col_names = FALSE, col_types = NULL, na = "", skip = 3)

challa <- rename(challa, `Date` = `...1`, `Varve Number`= `...2`, `VarveThickness` = `...3`, `LightThickness` = `...4`, `DarkThickness`= `...5`)

ggplot(data = challa, mapping = aes(x= Date))+
 geom_smooth(aes(y= LightThickness, color = "LightThickness"))+
  geom_smooth(aes(y=DarkThickness, color = "DarkThickness")) +
  geom_smooth(aes(y= VarveThickness, color="VarveThickness"))+
  labs(title = "Measurement of sediment deposit on the bed of Lake Challa", caption= "Varve is the cumulative value of DarkThickness and LightThickness.", fill= "Layers")+
  xlab("Date (AD)")+
  ylab("Thickness of layer in milimeters")
```


## Tanganyika (Trent)
```{r, message=FALSE}
VOI_Rainfall <- read_excel("East Africa lake data.xlsx", sheet = "Tanganyika")
```

## Malawi (Tyler)
```{r, message=FALSE, warning=FALSE}
malawi <- read_xlsx("East Africa lake data.xlsx", sheet = "Malawi", skip = 2, col_names = FALSE)

part1 <- malawi %>%
  select('...1', '...2', '...3')%>%
  rename(BSi = '...3', 'Calender Year' = '...2', 'Not Cal' = '...1')
part1$'Calender Year' <- round(part1$'Calender Year')

part2 <- malawi %>%
  select('...6', '...7', '...8')%>%
  rename('BSi 2' = '...8', 'Calender Year' = '...7', 'Cal' = '...6')
part2$'Calender Year' <- round(part2$'Calender Year')


myData <- full_join(part1,part2) %>%
  gather('Cal', 'Not Cal', value = 'Years Before 2002', key = 'LakeData') %>%
  gather(BSi, 'BSi 2', key = Delete, value = BSi) %>%
  mutate(Delete = LakeData) %>%
  filter(LakeData == Delete) %>%
  select(-Delete) %>%
  rename(Year = 'Calender Year') %>%
  filter(LakeData == 'Cal')

ggplot(data = myData) +
  geom_smooth(mapping = aes(x = Year, y = BSi), se = FALSE) +
  labs(title = "BSi index of Malawi Lake Data over Time") + 
  theme_dark()
```

## CRU and MAM Rainfall (Jack)
```{r, message=FALSE}
CRUnMAM_Rainfall <- read_excel("East Africa lake data.xlsx", sheet = 9)
CRUnMAM_Rainfall_tidy <- select(CRUnMAM_Rainfall, 'GPCC MAM Normalized',	'CRU MAM Normalized',	'GPCC OND Normalized', 'CRU OND Normalized')
CRUnMAM_Rainfall_tidy <- rename(CRUnMAM_Rainfall, GPCC_MAM = 'GPCC MAM Normalized',	CRU_MAM = 'CRU MAM Normalized',	GPCC_OND = 'GPCC OND Normalized', CRU_OND = 'CRU OND Normalized') %>% gather(GPCC_MAM, CRU_MAM,	GPCC_OND, CRU_OND, key = "Rainfall_type", value = "value")

ggplot(CRUnMAM_Rainfall_tidy, aes(y = value, x = Year, color = Rainfall_type)) +
geom_line()+
labs(title = "Rainfall Types for every year", x = "Year", y = "Rainfall")
```


## VOI Rainfall (Logan)
```{r, message=FALSE}
VOI_Rainfall <- read_excel("East Africa lake data.xlsx", sheet = "VOI rainfall station data")
VOI_Rainfall_tidy <- VOI_Rainfall %>% select(-c(`...14`, `...15`)) %>% 
  gather("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec", key = "Month", value = "Rainfall") %>%
  mutate(Month = if_else(Month == "Jan", 1, if_else(Month == "Feb", 2, if_else(Month == "Mar", 3, if_else(Month == "Apr", 4, if_else(Month == "May", 5, if_else(Month == "Jun", 6, if_else(Month == "Jul", 7, if_else(Month == "Aug", 8, if_else(Month == "Sep", 9, if_else(Month == "Oct", 10, if_else(Month == "Nov", 11, if_else(Month == "Dec", 12, 0))))))))))))) %>% 
  arrange(Year) %>%
  filter(!is.na(Rainfall))

labels <- c(Female = "Women", Male = "Men")

ggplot(VOI_Rainfall_tidy, aes(x = Year, y = Rainfall)) + 
  facet_wrap(.~Month, labeller=labeller(Month =  c(`1` = "January", `2` = "February", `3` = "March", `4` = "April", `5` = "May", `6` = "June", `7` = "July", `8` = "August", `9` = "September", `10` = "October", `11` = "November", `12` = "December"))) + 
  geom_line() + 
  labs(title = "VOI Rainfall for every month vs. year", y = "Rainfall (cm)", x = "Year")
```




