---
title: "Lab13"
author: "Spicy Bois"
date: "11/18/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(forcats)
library(gapminder)
```


## Individual Sections

### Tyler's Section
#### The Deaths in Moldova are much greater then the deaths of Austria
**DataSet:** The Dataset I use is the Total Deaths Dataset. This Dataset has all the deaths that have happened in the country for each year for every country starting in 1990 all the way to 2016. I also have a dataset that is the amount of deaths per 100000 people. This has the same stats as the total deaths table

```{r}

perm_mean <- function(perms = 1000, all_values, n_A)
{
  ## Variables ##
  # perms: The number of permutations 
  # all_values (num): all data values
  # n_A (int): Size of group A
  ###############
  
  # Step 1:
  # Create vector of zeroes of length "perms" to store
  # permuted mean differnces
  perm_mean_diffs <- numeric(perms)
  
  # Loop throught number of permutations
  for (i in c(1:perms))
  {
    # Step 2:
    # Randomly separate vector "values" into disjoint 
    # groups of size "nA" and "nB" respectively
    rand_order <- sample(all_values)
    
    # Step 3:
    # Compute the sample means for the two groups from 
    # step 2(
    groupA <- rand_order[1:n_A]
    groupB <- rand_order[(n_A+1):length(all_values)]
    
    # Step 4: 
    # Compute the difference in sample means, store the
    # value in the vector from step 1
    mean_dif<- mean(groupA) - mean(groupB)
    perm_mean_diffs[i] <- mean_dif
  }
  
  # Step 5:
  # Return the permuted mean differences vector
  return(perm_mean_diffs)
  
}

```


```{r}
deaths <- read_csv("murder_total_deaths.csv") 
deathTotal <- deaths %>% gather('1990':'2016', key = 'Year', value = 'Total_Deaths') %>% filter(country == 'Moldova' | country == 'Austria')
deathPerOnek <- read_csv("murder_per_100000_people.csv") %>% gather('1950':'2016', key = 'Year', value = 'Death_Ratio') %>% filter(country == 'Moldova' | country == 'Austria', Year >= 1990)

deathFacts <- full_join(deathTotal, deathPerOnek)
deathau <-deathFacts %>% filter(country == "Austria")
deathmold <-deathFacts %>% filter(country == "Moldova")
meanTest <- perm_mean(1000, deathFacts$Total_Deaths, 27)
meanTest2 <- perm_mean(1000, deathFacts$Death_Ratio, 27)

ggplot(data = deathFacts) +
  geom_point(mapping = aes(x = Year, y = Total_Deaths, color = country))+
  labs(title = "Total Deaths per Year", x = "Year", y = "Total Deaths")
  
ggplot(data = deathFacts) +
  geom_point(mapping = aes(x = Year, y = Death_Ratio, color = country))+
  labs(title = "Deaths per 100000 People Per Year", x = "Year", y = "Deaths per 100000 People")

ggplot() +
  geom_histogram(mapping = aes(meanTest)) +
  geom_vline(xintercept = (mean(deathmold$Total_Deaths) - mean(deathau$Total_Deaths)), color = "Red") +
  labs(title = "Histogram of the Mean Total Deaths of Moldova - Austria", x = "Mean of Deaths", y = "Count", caption = "Red Line is the True Data\nPercentile = 99.99001")

ggplot() +
  geom_histogram(mapping = aes(meanTest2)) +
  geom_vline(xintercept = (mean(deathmold$Death_Ratio) - mean(deathau$Death_Ratio)), color = "Red") +
  labs(title = "Histogram of the Mean Deaths per 100000 People of Moldova - Austria", x = "Mean of Deaths Per 100000 People", y = "Count", caption = "Red Line is the True Data\nPercentile = 99.99001")

map(deaths, mean)


```
**Permutation Test: ** My Null Hypothosis was that the mean deaths of Moldovo are the same as Austria. My Alternative Hypothosis was that Moldova's deaths were higher then Austria's. My conclusion is that we reject the null hypothesis with a percentile pf 99.99001. We have sufficient evidence that the mean deaths of Moldova is bigger then the mean deaths of Austria

**Map Function: ** I used the Map function to find the mean amount of deaths for each year using the dataset for total amount of deaths per year per country.


### Logan's Sections
#### Is there a correlation between energy production and gcp?
**DataSet:** The Dataset I use is the International Energy Satitics. This has all energy consumptions and production of the continents from 1980 to 2016. 

```{r}
perm_cor <- function(perms = 1000, x, y) {
  
  correlations <- vector("double", length = perms)
  
  for (i in c(1:perms)) {
    temp <- sample(y)
    
    correlations[i] <- cor(x, temp)
  }
  
  return(correlations)
}
```


```{r, warning=FALSE, message=FALSE}
energy_data <- read_csv("energy.csv") %>% 
  gather(`1980`, `1981`, `1982`, `1983`, `1984`, `1985`, `1986`, `1987`, `1988`, `1989`, `1990`, `1991`, `1992`, `1993`, `1994`, `1995`, `1996`, `1997`, `1998`, `1999`, `2000`, `2001`, `2002`, `2003`, `2004`, `2005`, `2006`, `2007`, `2008`, `2009`, `2010`, `2011`, `2012`, `2013`, `2014`, `2015`, `2016`, key = "year", value = "quadBtu") %>% 
  select(-c("unit")) %>% mutate(year = as.numeric(year))

continents = c("Africa", "Americas", "Asia & Oceania", "Europe")
gapminder_logan <- gapminder %>% mutate(continent = ifelse(continent %in% c("Asia", "Oceania"), "Asia & Oceania", continents[continent])) %>% group_by(continent, year) %>% summarise(lifeExp = mean(lifeExp), pop = mean(pop), gdpPercap = mean(gdpPercap)) %>% mutate(year = as.numeric(year))

joinedData <- inner_join(gapminder_logan, energy_data, by = c("continent", "year"))

ggplot(joinedData, aes(x = year)) + facet_wrap(.~continent) + geom_line(aes(y = quadBtu, color = type)) #+ geom_line(aes(y = gdpPercap))
ggplot(joinedData, aes(x = year)) + facet_wrap(.~continent) + geom_line(aes(y = gdpPercap))

filtered_joinedData <- joinedData %>% filter(type == "Production")

cor_test <- perm_cor(1000, filtered_joinedData$quadBtu, filtered_joinedData$gdpPercap)

ggplot() +
  geom_histogram(mapping = aes(cor_test)) +
  geom_vline(xintercept = cor(filtered_joinedData$quadBtu, filtered_joinedData$gdpPercap), color = "Red") +
  labs(title = "Histogram of the Correlation between GCP and Energy Production", x = "Correlation between GCP and Energy Production", y = "Count", caption = "Red Line is the True Data\nPercentile = 99.99001")

filtered_joinedData2 <- joinedData %>% filter(type == "Consumption")

cor_test <- perm_cor(1000, filtered_joinedData2$quadBtu, filtered_joinedData2$gdpPercap)

ggplot() +
  geom_histogram(mapping = aes(cor_test)) +
  geom_vline(xintercept = cor(filtered_joinedData2$quadBtu, filtered_joinedData2$gdpPercap), color = "Red") +
  labs(title = "Histogram of the Correlation between GCP and Energy Consumption", x = "Correlation between GCP and Energy Consumption", y = "Count", caption = "Red Line is the True Data\nPercentile = 99.99001")
```

**Conclusion:** Based on the resaults above, there does not seem to be any correlation between gdp and energy production. However, there is a high correlation between GCP and energy consumption, however the correlation is too small to make an difinite conclusion.



#Joe's Section
#Null Hypothesis: The largest countries in the world have the same polution per capita than smaller countries. My alternative hypothesis is that the largest countries have a higher polution per capita than small countries. 
#Datasets: I am using the dataset "CO2 emmissions (tonnes per person)" from Gapminder
```{r}
library(tidyverse)
library(readxl)
polperson <- read_csv('yearlyco2person.csv', check.names = FALSE)

poltop <- polperson %>% gather('2004':'2014', key ="Year", value = 'Total') %>% filter(country == 'United States' | country == "China" | country == "India") %>% select(Year, Total, country) %>% type.convert("Year", double) %>% type.convert("country", as.is = TRUE)
polbottom <- polperson %>% gather('2004':'2014', key ="Year", value = 'Total') %>% filter(country == "Tuvalu" | country == "Nauru" | country == "Palau") %>% select(Year, Total, country) %>% type.convert("Year", double) %>% type.convert("country", as.is = TRUE)
topmean <- polperson %>% gather('2004':'2014', key ="Year", value = 'Total') %>% select(Year, country, Total) %>% filter(country == 'United States' | country == "China" | country == "India") %>% type.convert("Year", double) %>% select(Total)
botmean <- polperson %>% gather('2004':'2014', key ="Year", value = 'Total') %>% select(Year, country, Total) %>% filter(country == "Tuvalu" | country == "Nauru" | country == "Palau") %>% select(Year, Total, country) %>% type.convert("Year", double) %>% select(Total)

A <- map_dbl(topmean, mean) #grabbing a mean value for CO2 output of largest countries from 2004 to 2014
B <- map_dbl(botmean, mean) #grabbing a mean value for CO2 output of smallest countries from 2004 to 2014

ggplot(data = poltop)+
 geom_smooth(mapping = aes(x= Year , y = Total, color = country), se = FALSE)+
  labs(title = "Yearly Carbon Output of Most Populous Countries per Person", x = "Year", y = "Output in thousands of tonnes per person") +
  geom_hline(yintercept = A)+
  geom_hline(yintercept = B)

ggplot(data = polbottom)+
 geom_smooth(mapping = aes(x= Year , y = Total, color = country), se = FALSE)+
  labs(title = "Yearly Carbon Output of Least Populous Countries per Person", x = "Year", y = "Output in thousands of tonnes per person") +
  geom_hline(yintercept = B)+
  geom_hline(yintercept = A)
```
#I wanted to know whether the industry of a populous country offset the inefficiencies of a relatively unpopulated country. I used the map function just to grab a quick mean of the 3 top and 3 bottom countries to compare the two sets of countries. I found that the mean values of the top 3 countries was far greater than that of the bottom 3 countries. I found that the ALTERNATIVE HYPOTHESIS was infact true and the null hypothesis was not. I attempted to integrate the perm_mean that I developed with Tyler, but I could not do so without making my code unfunctional.


## Trent Collins Indivdual

### Question: Is there a correlation between total energy production and the total number of people effected (injured/killed) by storms in the United States?

**Null Hypothesis: ** There is 0 correlation between total energy produced and people effected by dangerous storms.

**Alternative Hypothesis: ** There is non-zero, positive correlation between total energy produced and people effected by dangerous storms.

The test statistic that I am using is correlation.

```{r, echo=FALSE, messages = FALSE}
library(tidyverse)
library(gapminder)
library(readr)
library(dplyr)
library(stringr)

storm <- read_csv("storm_affected_annual_number.csv")
energy <- read_csv("energy_production_total.csv")

us_Storm <- storm %>%
  filter(country == "United States") %>%
  select("1970":"2008") %>% 
  gather("1970":"2008", key = "year", value = "People effected by storms")

us_Energy <- energy %>%
  filter(country == "United States") %>%
  select("1970":"2008") %>% 
  gather("1970":"2008", key = "year", value = "Total energy produced")

us_Data <- us_Storm %>% left_join(us_Energy, by = "year") %>% filter(!is.na("People effected by storms")) %>% filter(!is.na("Total energy produced"))

perm_cor <- function(perms = 1000, x, y)
{
  vec <- numeric(perms)
  for (i in c(1:perms))
  {
    r_Y <- sample(y)
    vec[i] <- cor(x, r_Y)
  }
  return(vec)
}

justStorm <- c(us_Data$`People effected by storms`)
justStorm[is.na(justStorm)] <- 0

justEnergy <- c(us_Data$`Total energy produced`)

us_data_corr <- perm_cor(1000, justEnergy, justStorm)
actual_corr <- cor(justEnergy, justStorm)


ggplot() + 
  aes(us_data_corr)+ 
  geom_histogram(binwidth=.05, colour="black", fill="grey") +   geom_vline(xintercept = actual_corr, color="red") + labs(title = "Correlation between Energy Production and the Severity of Storms in the USA", x ="correlation")

p_value <- 1 - sum(us_data_corr < actual_corr)/1000 #.012
p_value

my_percentile <- sum(us_data_corr <   actual_corr)/length(us_data_corr) *100 #98.8
my_percentile
```
**Analysis: ** The real data value falls in the 99th percentile of the data, and the p value of the real data is .014, indicating that the real data has a statistically significant correlation. Due to the low p-value, I reject the null hypothesis. This indicates that there is a correlation between severe storms and higher energy production.

**Speculation: ** These findings make sense, as the higher energy production and consumption has been linked to recent climate change and global warming. Warmer air can contain more water vapor than cooler air. Global analyses show that the amount of water vapor in the atmosphere has in fact increased due to human-caused warming. This extra moisture is available to storm systems, resulting in heavier rainfalls, and more intense storms, which can cause more casualties and injuries.
