---
title: "Lab12"
author: "Spicy Bois"
date: "11/13/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)

library(tidyverse)
library(forcats)
library(gapminder)

mydata <- gapminder_unfiltered
```

# Gapminder Questions

## Your data science team must come up with two specific questions of interest about the gapminder data set (from 'gapminder' library). 

1) What is the difference between the life expectancy between China and Great Britan?

2) Is there a corrilation between life expectancy and gpd?

## Briefly describe the gapminder dataset.

The gapminder is a dataset that charts data such as life exp, population, and gdp for most countries in the world. This data is recodered for each year form 1950 to 2007.

## Explain why your questions are important. One of the questions should be answered with the two-sample mean permutation test, and the other with the correlation permutation test.

These questions are important in order to figure out how to best increase the life expentancy of a certian country. THis is important because everyone wants to life longer.

## Finish the function skeletons to implement the two-sample permuation test and the correlation permutation test. Run a two-sample mean permutation test. Run a correlation permutation test.

### Two-sample permuation test 

Perm_mean creates 1000 iterations of the two-sample permuation test. Fill in the function-skeleton to implement the function.

```{r, eval = F}
perm_mean <- function(perms = 1000, all_values, n_A) {
  ## Variables ##
  # perms: The number of permutations 
  # all_values (num): all data values
  # n_A (int): Size of group A
  ###############
  
  # Step 1:
  # Create vector of zeroes of length "perms" to store
  # permuted mean differnces
  perm_mean_diffs <- numeric(perms)
  
  # Loop throught number of permutations
  for(i in c(1:perms)) {
    # Step 2:
    # Randomly separate vector "values" into disjoint 
    # groups of size "nA" and "nB" respectively
    rand_order <- sample(all_values)
    group_A <- rand_order[1:n_A]
    group_B <- rand_order[(n_A+1):length(all_values)]
    
    # Step 3:
    # Compute the sample means for the two groups from 
    # step 2
    A_bar <- mean(group_A)
    B_bar <- mean(group_B)
    
    
    # Step 4: 
    # Compute the difference in sample means, store the
    # value in the vector from step 1
    diff_means <- A_bar - B_bar
    perm_mean_diffs[i] <- diff_means
    
  }
  
  # Step 5:
  # Return the permuted mean differences vector
  return(perm_mean_diffs)
  
}
```

### Linear Correlation Permuation Test

Perm_cor will create 1000 iterations of the correlation permuation test. Fill in the function-skeleton to implement the function.

```{r, eval = F}
perm_cor <- function(perms = 1000, x, y) {
  ## Variables ##
  # perms: The number of permutations 
  # x: Vector of Variable 1 - for computing correlation
  # y: Vector of Variable 2 - for computing correlation
  ###############
  
  # Step 1:
  # Create vector of zeroes of length "perms" to store
  # permuted mean differnces
  corrs <- numeric(perms)
  
  # Loop throught number of permutations
  for (i in c(1:perms)) {
    # Step 2:
    # Randomly mix up the values in the vector "y"
    randY <- sample(y)
    
    # Step 3:
    # Compute the correlation between x and the randomly mixed
    # up y-vector. Store this value in the vector from step 1.
    corrs[i] <- cor(x, randY)
  }
  
  # Step 4:
  # Return new updated vector, created in step 1
  return(corrs)
  
}
```

## Create visualizations of the statistical tests.

```{r}
#ggplot(mydata, aes(x = perm_mean(3313, mydata$lifeExp, 2))) + geom_histogram()
```

```{r}
#ggplot(mydata, aes(x = perm_cor(3313, mydata$lifeExp, mydata$pop))) + geom_histogram()
```

## Make data-based conclusions about the results. Do you believe there are differences among the groups you are comparing? 

## Generate percentiles, summary statistics (max, min), and probabilites (p-values) to justify your conclusions.

## Team Report

**Tyler:** I made the perm_cor function. I did this by first making a correlation vector full of 0's based on the number of inputted perms. Then I go into the for loop and in the for loop I first shuffle the Y vector and then find the correlation of X and the shuffled Y. I then take that correlation value and store it into spot "i" of the vector of correlations. I then return the vector of correlations

**Trent:** 

**Logan:** 

**Joe**  I assisted creating the perm_cor function. I also answered a couple misc. questions throughout the lab.
