---
title: "Lab12"
author: "Spicy Bois"
date: "11/13/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
```

# Lab 12

#Two-sample permuation test 

#The Gapminder dataset contains data that is used to describe the size, wealth, and health of all nations in the world over a time period of 1952 to 2007.

Perm_mean creates 1000 iterations of the two-sample permuation test. Fill in the function-skeleton to implement the function.

```{r, eval = F}
library(tidyverse)
library(forcats)
library(gapminder)
perm_mean <- function(perms = 1000, all_values, n_A) {
  ## Variables ##
  # perms: The number of permutations 
  # all_values (num): all data values
  # n_A (int): Size of group A
  ###############
  
  # Step 1:
  # Create vector of zeroes of length "perms" to store
  # permuted mean differnces
  perm_mean_diffs <- numeric(perms)
  
  # Loop throught number of permutations
  for(i in c(1:perms)) {
    # Step 2:
    # Randomly separate vector "values" into disjoint 
    # groups of size "nA" and "nB" respectively
    rand_order <- sample(all_values)
    group_A <- rand_order[1:n_A]
    group_B <- rand_order[(n_A+1):length(all_values)]
    
    # Step 3:
    # Compute the sample means for the two groups from 
    # step 2
    A_bar <- mean(group_A)
    B_bar <- mean(group_B)
    
    
    # Step 4: 
    # Compute the difference in sample means, store the
    # value in the vector from step 1
    diff_means <- A_bar - B_bar
    perm_mean_diffs[i] <- diff_means
    
  }
  
  # Step 5:
  # Return the permuted mean differences vector
  return(perm_mean_diffs)
  
}
```

# Linear Correlation Permuation Test

Perm_cor will create 1000 iterations of the correlation permuation test. Fill in the function-skeleton to implement the function.

```{r, eval = F}

perm_cor <- function(perms = 1000, x, y)
{
  ## Variables ##
  # perms: The number of permutations 
  # x: Vector of Variable 1 - for computing correlation
  # y: Vector of Variable 2 - for computing correlation
  ###############
  
  # Step 1:
  # Create vector of zeroes of length "perms" to store
  # permuted mean differnces
  corrs<-numeric(perms)
  
  # Loop throught number of permutations
  for (i in c(1:perms))
  {
    # Step 2:
    # Randomly mix up the values in the vector "y"
    randY<-sample(y)
    # Step 3:
    # Compute the correlation between x and the randomly mixed
    # up y-vector. Store this value in the vector from step 1.
    corrs[i]<-cor(x,randY)
    
  }
  
  # Step 4:
  # Return new updated vector, created in step 1
  return(corrs)
  
}
```

# Gapminder Questions

Your data science team must come up with two specific questions of interest about the gapminder data set (from 'gapminder' library). 

* Briefly describe the gapminder dataset.

* Explain why your questions are important. One of the questions should be answered with the two-sample mean permutation test, and the other with the correlation permutation test.

* Finish the function skeletons to implement the two-sample permuation test and the correlation permutation test. Run a two-sample mean permutation test. Run a correlation permutation test.

* Create visualizations of the statistical tests.

* Make data-based conclusions about the results. Do you believe there are differences among the groups you are comparing? 

* Generate percentiles, summary statistics (max, min), and probabilites (p-values) to justify your conclusions.

* Summarize each individual's contribution to the lab.

##Team Report

**Tyler:** I made the perm_cor function. I did this by first making a correlation vector full of 0's based on the number of inputted perms. Then I go into the for loop and in the for loop I first shuffle the Y vector and then find the correlation of X and the shuffled Y. I then take that correlation value and store it into spot "i" of the vector of correlations. I then return the vector of correlations

**Trent:** 

**Joe**  I assisted creating the perm_cor function. I also answered a couple misc. questions throughout the lab.
